### 2012-02-23 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

\FORALL{$stepNode \in pathType.stepNodes$}
	\IF{$stepNode.isAxisStep()$}
		\STATE $axisKind = stepNode.getAxisNode().getAxisKind()$
		\IF{$axisKind$ $!=$ $CHILD$ $AND$ $axisKind$ $!=$ $DESCENDANT$}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDFOR

\STATE $detailNode = stepNode.getChild(ExprNode)$
\IF{$detailNode != null$}
	\IF{$is(detailNode, VarRefNode)$}
		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) ==$ false}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDIF

@@

\begin{algorithm}
\caption{Function isWithoutPredicates}
\label{ALG_function_is_without_predicates}
\begin{algorithmic}[1]
\REQUIRE $pathType$: A $PathType$ instance.
\ENSURE A boolean result.

\FORALL{$stepNode \in pathType.stepNodes$}
	\IF{$stepNode.hasPredicates()$}
		\RETURN false
	\ENDIF
	\IF{$pathType.forBoundSubsteps[stepNode]$ $!=$ $null$}
		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode]$ $==$ false}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDFOR

@@

Function $endsWithExactlyOnePredicate$ checks if a path represented by a given $PathType$ instance contains exactly one predicate in its last step and if so, the predicate is returned. Otherwise, it returns false.

@@

\STATE $lastStepNode =$ last item from $pathType.stepNodes$
\IF{$lastStepNode.hasPredicates()$ $==$ false}
	\RETURN false
\ENDIF
\IF{number of item in list $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
	\RETURN false
\ENDIF

@@

@@

\IF{$is(node, FunctionCallNode)$}
    \IF{$node.fncName$ is one of $distinct-values$, $min$, $max$, $sum$}
        \STATE $argument = node.getChild(0)$
        \IF{$getType(argument) == PathType$ using only child and descendant axes, without predicates except the last step}
            \STATE memorize $argument$ as the negative key statement, weight 1
        \ENDIF
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
        \IF{$is(bindingNode, ForClauseNode)$}
            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
            \IF{$getType(expr) == PathType$ using only child and descendant axes}
                \STATE add $bindingNode.varName$ to $forVars$
            \ENDIF
        \ENDIF
    \ENDFOR

@@

\IF{$is(exprNode, OperatorNode$)}
    \IF{$exprNode.operator == GEN\_EQUALS$}
        \FORALL{$var \in forVars$}
            \IF{$exprNode$ is of a form $var/L = C$ where $L$ is a path without predicates, except the last step, using only child and descendant axes, and C is a literal constant}
                \STATE memorize $var/L$ as the negative key statement, weight 0.9
            \ENDIF
        \ENDFOR
    \ELSIF{$exprNode.operator == AND$}
        \STATE $processWhere(forVars, exprNode.leftSide)$
        \STATE $processWhere(forVars, exprNode.rightSide)$
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

>>>>

\FORALL{$stepNode \in pathType.stepNodes$}
	\IF{$stepNode.isAxisStep()$}
		\STATE $axisKind$ = $stepNode.getAxisNode().getAxisKind()$
		\IF{$axisKind$ $\neq$ $CHILD$ AND $axisKind$ $\neq$ $DESCENDANT$}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDFOR

\STATE $detailNode$ = $stepNode.getChild(ExprNode)$
\IF{$detailNode \neq$ null}
	\IF{$is(detailNode, VarRefNode)$}
		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode])$ == false}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDIF

@@

\begin{algorithm}
\caption{Function isWithoutPredicates}
\label{ALG_function_is_without_predicates}
\begin{algorithmic}[1]
\REQUIRE $pathType$: A $PathType$ instance.
\ENSURE A boolean result.getTargetReturnPathTypesRecursive

\FORALL{$stepNode \in pathType.stepNodes$}
	\IF{$stepNode.hasPredicates()$}
		\RETURN false
	\ENDIF
	\IF{$pathType.forBoundSubsteps[stepNode]$ $\neq$ null}
		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode]$ == false}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDFOR

@@

Function $endsWithExactlyOnePredicate$ in Algorithm \ref{ALG_function_ends_with_exactly_one_predicate} checks if a path represented by a given $PathType$ instance contains exactly one predicate in its last step and if so, the predicate is returned. Otherwise, it returns false.

@@

\STATE $lastStepNode$ = last item from $pathType.stepNodes$
\IF{$lastStepNode.hasPredicates()$ $==$ false}
	\RETURN false
\ENDIF
\IF{number of item in list $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
	\RETURN false
\ENDIF

@@

Function $isTargetPath$ in Algorithm \ref{ALG_function_is_target_path} checks if a path represented by a given $PathType$ instance is so-called target path for the specified variable. A target path for a variable is a path where its first step is a reference to the variable.

\begin{algorithm}
\caption{Function isTargetPath}
\label{ALG_function_is_target_path}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $pathType$: A $PathType$ instance. \\
  $varName$: A variable name.
}
\ENSURE Returns true, if the first step of path represented by the $PathType$ instance is a reference to the specified variable.

\todo[inline]{Upravit styl vsetkych algoritmov (null, ==, !=, true/false, ...) tak ako je tu.}

\STATE $firstStep$ = first item from  $pathType.stepNodes$
\STATE $detailNode$ = $firstStep.detailNode$
\IF{$detailNode$ == null}
	\RETURN false
\ENDIF
\IF{$is(detailNode, VarRefNode)$ == false}
	\RETURN false
\ENDIF
\IF{$detailNode.varName$ == $varName$}
	\RETURN true
\ELSE
	\RETURN false
\ENDIF
\end{algorithmic}
\end{algorithm}

Function $getTargetReturnPathTypes$ in Algorithm \ref{ALG_function_get_target_return_path_types} searches the given FLWOR for $PathTypes$ representing so-called target return paths. A target return path is a return path in a where clause of a FLWOR. The FLWOR is processed recursively using function $getTargetReturnPathTypesRecursive$ in Algorithm \ref{ALG_function_get_target_return_path_types_recursive}.

\begin{algorithm}
\caption{Function getTargetReturnPathTypes}
\label{ALG_function_get_target_return_path_types}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $flworNode$: A $FLWORExprNode$ instance to search for target return paths. \\
  $varName$: A variable name.
}
\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by $PathType$ instances.

\STATE $pathTypes$ = an empty array
\STATE $getTargetReturnPathTypesRecursive(flworNode.$ $getChild(ReturnClauseNode), varName, pathTypes)$
\RETURN $pathTypes$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Function getTargetReturnPathTypesRecursive}
\label{ALG_function_get_target_return_path_types_recursive}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $node$: A node to search for target return paths. \\
  $varName$: A variable name. \\
  $pathTypes$ An output list to add found target return path types to.
}

\IF{$is(node, ExprNode)$}
	\STATE $type$ = $getType(node)$
	\IF{$type$ is $PathType$ AND $isTargetPath(type, varName)$}
		\STATE add $type$ to $pathTypes$
	\ENDIF
\ENDIF
\FORALL{$child \in node.getChildren()$}
	\STATE $getTargetReturnPathTypesRecursive(child, varName, pathTypes)$
\ENDFOR
\end{algorithmic}
\end{algorithm}

A similar couple of functions is function $getTargetReturnPaths$ and function $getTargetReturnPathsRecursive$. Only difference between those two couples of functions is that the former one searches for all expressions which are of $PathType$ representing a target return path, while the later one searches only for instances of $PathExprNode$.

To ilustrate this difference, we introduce the following example. Assuming $P$ is a return target path expression for variable $var$ and $flwor$ is a FLWOR expression with return clause $data(P)$, function call $getTargetReturnPathTypes(flwor, var)$ returns two $PathTypes$. One is the type of expression $P$ and the other is the type of the function call $data(P)$ which is also $PathType$. On the other hand, function call $getTargetReturnPaths(flwor, var)$ returns just one $PathType$ representing $P$.

\begin{algorithm}
\caption{Function getTargetReturnPaths}
\label{ALG_function_get_target_return_paths}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $flworNode$: A $FLWORExprNode$ instance to search for target return paths. \\
  $varName$: A variable name.
}
\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by $PartExprNode$ instances.

\STATE $paths$ = an empty array
\STATE $getTargetReturnPathsRecursive(flworNode.$ $getChild(ReturnClauseNode), varName, paths)$
\RETURN $paths$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Function getTargetReturnPathsRecursive}
\label{ALG_function_get_target_return_paths_recursive}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $node$: A node to search for target return paths. \\
  $varName$: A variable name. \\
  $paths$ An output list to add found target return path types to.
}

\IF{$is(node, PathExprNode)$}
	\STATE $type$ = $getType(node)$
	\IF{$type$ is $PathType$ AND $isTargetPath(type, varName)$}
		\STATE add $type$ to $paths$
	\ENDIF
\ENDIF
\FORALL{$child \in node.getChildren()$}
	\STATE $getTargetReturnPathsRecursive(child, varName, paths)$
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

\IF{$is(node, FunctionCallNode)$}
    \IF{$node.fncName$ is one of $distinct-values$, $min$, $max$, $sum$}
        \STATE $argument = node.getChild(0)$
        \STATE $type = getType(argument)$
        \IF{$type$ is $PathType$ AND $usesOnlyChildAndDescendantAxes(type)$ AND $isWithoutPredicatesExceptLastStep(type)$}
            \STATE memorize $argument$ as the negative key statement, weight 1
        \ENDIF
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
        \IF{$is(bindingNode, ForClauseNode)$}
            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
            \IF{$getType(expr)$ is $PathType$ AND $usesOnlyChildAndDescendantAxes(getType(expr))$}
                \STATE add $bindingNode.varName$ to $forVars$
            \ENDIF
        \ENDIF
    \ENDFOR

@@

\IF{$is(exprNode, OperatorNode$)}
    \IF{$exprNode.operator == GEN\_EQUALS$}
        \FORALL{$var \in forVars$}
            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is $PathType$ AND $isWithoutPredicatesExceptLastStep(getType(var/L))$ AND $usesOnlyChildAndDescendantAxes(getType(var/L))$ AND $C$ is a literal constant}
                \STATE memorize $var/L$ as the negative key statement, weight 0.9
            \ENDIF
        \ENDFOR
    \ELSIF{$exprNode.operator == AND$}
        \STATE $processWhere(forVars, exprNode.leftSide)$
        \STATE $processWhere(forVars, exprNode.rightSide)$
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}
